---
title: "Mixed-Effects Analysis: Transforms vs Baseline"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(performance)
library(broom.mixed)
# Optional: install.packages("glmmTMB") if you haven't
# library(glmmTMB)
````

## 1. Data import

```{r import}
# Replace with your actual data paths or loading code

# Example:
# df      <- read_csv("df.csv")      # has transform_type, grid_size, complexity, instance_id, ..., error_rate
# df_base <- read_csv("df_base.csv") # same minus transform_type (baseline condition)

# If df and df_base are already in the environment from RStudio, this chunk can be left as-is.
```

## 2. Combine baseline and transform data

```{r combine}
df_base2 <- df_base %>%
  mutate(transform_type = "baseline")

df_all <- bind_rows(df, df_base2) %>%
  mutate(
    transform_type = factor(transform_type),
    grid_size      = factor(grid_size),
    complexity     = factor(complexity)
  ) %>%
  droplevels()

# Set baseline as reference
df_all <- df_all %>%
  mutate(transform_type = relevel(transform_type, ref = "baseline"))

# Quick check
df_all %>% count(transform_type)
```

## 3. Exploratory plots

```{r eda-overall, fig.width=7, fig.height=4}
ggplot(df_all, aes(x = transform_type, y = error_rate)) +
  geom_violin(fill = "grey90", color = "grey40") +
  geom_boxplot(width = 0.15, outlier.shape = NA) +
  labs(x = NULL, y = "error_rate", title = "Error rate by transform_type")
```

```{r eda-facets, fig.width=10, fig.height=7}
ggplot(df_all, aes(x = transform_type, y = error_rate)) +
  geom_boxplot(outlier.alpha = 0.3) +
  facet_grid(grid_size ~ complexity, labeller = label_both) +
  labs(x = NULL, y = "error_rate",
       title = "Error rate by transform_type within each (grid_size, complexity)")
```

## 4. Linear mixed-effects model

Model:
`error_rate ~ transform_type * grid_size * complexity + (1 | instance_id)`

```{r lmm-fit, message=FALSE, warning=FALSE}
lmm <- lmer(
  error_rate ~ transform_type * grid_size * complexity + (1 | instance_id),
  data = df_all,
  REML = FALSE
)

summary(lmm)
anova(lmm, type = 2)
```

## 5. Post-hoc contrasts vs baseline

### 5.1 Overall contrasts (averaged over grid_size & complexity)

```{r posthoc-main, message=FALSE, warning=FALSE}
emm_main <- emmeans(lmm, ~ transform_type)
contrast_main <- contrast(emm_main, method = "trt.vs.ctrl", ref = "baseline", adjust = "holm")
summary(contrast_main, infer = TRUE)
```

### 5.2 Contrasts by grid_size

```{r posthoc-grid, message=FALSE, warning=FALSE}
emm_grid <- emmeans(lmm, ~ transform_type | grid_size)
contrast_grid <- contrast(emm_grid, method = "trt.vs.ctrl", ref = "baseline", adjust = "holm")
summary(contrast_grid, infer = TRUE)
```

### 5.3 Contrasts by complexity

```{r posthoc-complexity, message=FALSE, warning=FALSE}
emm_cplx <- emmeans(lmm, ~ transform_type | complexity)
contrast_cplx <- contrast(emm_cplx, method = "trt.vs.ctrl", ref = "baseline", adjust = "holm")
summary(contrast_cplx, infer = TRUE)
```

## 6. Diagnostics

```{r diagnostics, fig.width=7, fig.height=6}
check_model(lmm)
```

```{r diagnostic-plots, fig.width=7, fig.height=6}
par(mfrow = c(1, 2))
plot(lmm, which = 1)  # residuals vs fitted
qqnorm(residuals(lmm)); qqline(residuals(lmm))
par(mfrow = c(1, 1))
```

## 7. Optional: Beta regression with random effects

(Use if `error_rate` is in (0,1) and you want a beta model instead of Gaussian.)

```{r beta-regression, eval=FALSE}
library(glmmTMB)

# Small adjustment to keep (0,1); tweak eps as needed
eps <- 1e-4
df_all_beta <- df_all %>%
  mutate(error_rate_beta = pmin(pmax(error_rate, eps), 1 - eps))

# Fit beta regression with random intercept
tmb <- glmmTMB(
  error_rate_beta ~ transform_type * grid_size * complexity + (1 | instance_id),
  data = df_all_beta,
  family = beta_family(link = "logit")
)

summary(tmb)

# Post-hoc on the response scale
emm_beta <- emmeans(tmb, ~ transform_type, type = "response")
contrast_beta <- contrast(emm_beta, method = "trt.vs.ctrl", ref = "baseline", adjust = "holm")
summary(contrast_beta, infer = TRUE)
```

## 8. Save model results

```{r save-results}
fixed_tbl <- tidy(lmm, effects = "fixed")
write_csv(fixed_tbl, "lmm_fixed_effects.csv")
saveRDS(lmm, "lmm_fit.rds")

fixed_tbl
```

```

That’s the whole pipeline: combine data → explore → fit LMM → contrasts vs baseline → diagnostics → optional beta model → save outputs. Once you knit this, you’ll have a reproducible analysis you can poke at alongside your plots.
```
